Lattice Boltzmann simulations of magnetically responsive\\ anisotropic particle stabilized emulsions

### 2025-02-16 chapters/analysis_techniques.tex
Changes 1. Added more analysis to application of magnetic fields during formation chapter 2. Added references to fix some (not all) literature reference issues 3. Added some detail in analysis techniques section

\subsection{Domain size}

@@

\subsection{Tortuosity}

@@

Before using taufactor, the microstructure is first binarized. This is done by first calculating the order parameter of the system with $\phi = \rho_b - \rho_r$ where $\phi$ is the order parameter and $\rho_b, \rho_r$ are the densities of the blue and red fluid respectively. Next, the interface described with $\phi$ is smoothened using the algorithm detailed in section \ref{section:filling_routine}. In many of the bijels synthesized using small molecule liquids, monomer is introduced into the oil phase and polymerized. Therefore, to perform the binarization step, we select all points where $\phi \leq 0$ and set their value to $0$, and all other values are set to $1$. To taufactor, $1$ represents void spaces and 0's are solid.

\subsection{Nematic order parameter}

@@

\subsection{Radial distribution function}

The radial distribution function (RDF) is used to measure the normalized number density of particles around a reference particle as a function of distance. The normalization factor is the expected density of particles derived from an ideal gas. The general form of the equation to calculate the RDF is $g(r) = \frac{1}{\rho_{ideal}(r) 4\pi r^2} \int_{r}^{r+\Delta r} \delta(r_{i} - r_{j}) dr$ where $\rho_{ideal}(r)$ represents the density expected at distance r from a reference particle in an ideal gas, $\delta(r_{i} - r_{j})$ represents the number of particles within the bin $r + \Delta r$ and integrated to get a volumetric density. As implemented in a numerical method, the integral becomes a sum. After correcting for the periodic boundary condition using the minimum image convention, the RDF can be used to characterize periodic systems. We utilize the implementation in the software package Freud.

\subsection{Average interface angle and proportion of particles on the interface}

@@

@@

\subsection{Yield stress and zero shear rate viscosity}

@@

>>>>

\subsection{Domain size}
\label{section:domain_size}

@@

\subsection{Curvature}
\label{section:curvature}

The simulation data was first pre-processed by calculating the order parameter of the densities, defined as the difference between the blue and red 
fluid densities at each grid point. $\phi(\vec{x}) = \rho^b(\vec{x}) - \rho^r(\vec{x})$. We then apply the particle filling algorithm described in 
section \ref{section:filling_routine} to the $\phi$ distribution, This reduces error from the presence of particles at the interface, 
resulting in an interface geometry that is more accurate to the bijel interface profile. From the filled order parameter, $\phi$, we calculate the mesh of 
the iso-surface representing the interface at $\phi = 0$ using a marching cubes method as implemented in scikit-image. We then use pyvista to calculate the 
gaussian and mean curvatures of the bijel microstructure, in addition to the interface area. The radii of curvature, $R_1, R_2$ are calculated from the 
mesh through measuring the angles between nearest neighbors and the laplacian of the faces to calculate the Gaussian and mean curvature respectively.

We then limit the values of the calculated Gaussian and mean curvatures to reduce the effect of numerical artefacts created during the meshing process. We 
limit the Gaussian and mean curvatures to the expected curvatures around $|H| \approx \frac{2}{L}$ and $|K| \approx (\frac{2}{L})^2$. $L$ here is defined 
as the characteristic length scale obtained from the first moment of the spherically averaged structure factor described in section \ref{section:domain_size}.  
We then calculate an interface area normalized curvatures to remove the differences in absolute curvature values arising from the change in the characteristic 
length scale by calculating the surface area to volume ratio, $\Sigma = A_{int}/V$ where $A_{int}$ is the interfacial area and $V$ is the system volume. After 
dividing the obtained curvatures with $Sigma$, we obtain the discrete quantities, $H \Sigma^{-1}$ and $K \Sigma^{-2}$. For data presented as a time or magnetic 
field series, we sum each curvature distributions yielding, $\langle H \Sigma^{-1} \rangle$ and $\langle K \Sigma^{-2} \rangle$. For a bijel with volume ratio 
$50:50$ as used here, we expect $\langle K \Sigma^{-2} \rangle < 0$ and $\langle H \Sigma^{-1} \rangle \approx 0$.

\subsection{Tortuosity}
\label{section:tortuosity}

@@

Before using taufactor, the microstructure is first binarized. This is done by first calculating the order parameter of 
the system with $\phi = \rho_b - \rho_r$ where $\phi$ is the order parameter and $\rho_b, \rho_r$ are the densities of 
the blue and red fluid respectively. Next, the interface described with $\phi$ is smoothened using the algorithm detailed 
in section \ref{section:filling_routine}. \texttt{taufactor} determines the tortuosity by comparison
of steady-state diffusive flow through a porous medium with the bulk
diffusive flow in a control volume of the same size, molecular
diffusivity, and driving force \cite{cooper_taufactor_2016}. We used
the \texttt{PeriodicSolver} to employ periodic boundary conditions. In many of the bijels synthesized using small molecule liquids, monomer is 
introduced into the oil phase and polymerized. Therefore, to perform the binarization step, we select all points where 
$\phi \leq 0$ and set their value to $0$, and all other values are set to $1$. To taufactor, $1$ represents void spaces 
and 0's are solid. \cite{cooper_taufactor_2016}

\subsection{Nematic order parameter}
\label{section:nematic_order_parameter}

@@

\subsection{Radial distribution function}
\label{section:radial_distribution}

The radial distribution function (RDF) is used to measure the normalized number density of particles around a 
reference particle as a function of distance. The normalization factor is the expected density of particles derived 
from an ideal gas. The general form of the equation to calculate the RDF is 
$g(r) = \frac{1}{\rho_{ideal}(r) 4\pi r^2} \int_{r}^{r+\Delta r} \delta(r_{i} - r_{j}) dr$ where $\rho_{ideal}(r)$ 
represents the density expected at distance r from a reference particle in an ideal gas, $\delta(r_{i} - r_{j})$ 
represents the number of particles within the bin $r + \Delta r$ and integrated to get a volumetric density. 
As implemented in a numerical method, the integral becomes a sum. After correcting for the periodic boundary 
condition using the minimum image convention, the RDF can be used to characterize periodic systems. We utilize 
the implementation in the software package Freud. \cite{ramasubramani_freud_2020}

\begin{equation}
    g(r) = \frac{n_p-1}{n_p} V \left\langle\delta\left(r-r_i\right)\right\rangle ,
\end{equation}

where \(n_p\) is the number of particles in the volume
\(V\), and \(\langle\cdot\rangle\) denotes an average over all
particles. The radial distribution function was calculated by binning
the distance between all particle pairs and normalizing the shells with
respect to the distribution \(4\pi \rho r^2 \mathrm{d}r\) of an ideal
gas.

\subsection{Average interface angle and proportion of particles on the interface}
\label{section:interface_angle}

@@

\subsection{Steinhardt order parameter}
\label{section:steinhardt_order_parameter}

Another factor affected by the application of the magnetic field is the Steinhardt 6 fold order parameter, $Q6$, which characterized the local 
particle ordering. \cite{steinhardt_bond-orientational_1983} It has been used to characterize crystallization, glass transitions and crystal 
structures for colloidal and molecular crystals. We utilize the average across all particles on the interface, $\langle Q6 \rangle$ and track 
this parameter as a function of time. Past work has identified that local crystallization of colloidal crystals takes place at $\langle Q6 \rangle = 0.38$, 
and that $\langle Q6 \rangle^{HCP} = 0.485$. \cite{steinhardt_bond-orientational_1983, toxvaerd_role_2020, mickel_shortcomings_2013} We utilize a voronoi 
cell derived definition of the neighbor list of each particle, addressing some of the shortcomings highlighted in the original distance based definition. 
\cite{steinhardt_bond-orientational_1983, mickel_shortcomings_2013}. We first define a complex orientational vector, $q_{lm}$

\begin{equation}
q_{lm}(i) = \frac{1}{N_b(i)} \Sigma_{j = 1}^{N_b(i)} Y_{lm}(\vec{R}_{ij})
\end{equation}

Where $N_b$ is the number of nearest neighbors of particle $i$, $l$ is controls to the degree of spherical harmonic to be calculated, $m$ is an integer 
defined as $-l \leq m \leq l$ that defines which spherical harmonic is being calculated and $\vec{r}_{ij}$ is the distance between two particle centers. 
The neighbor list $N_b$ is usually defined by a cutoff distance from the center of particle $i$ to its nearest neighbors $j$. Steinhardt then defined a 
bond order parameter $q_l$ from $q_{lm}$ as

\begin{equation}
Q_{l}(i) = \sqrt{\frac{4 \pi}{2l + 1} \Sigma_{m = -l}^{l} |q_{lm}(i)|^2}
\end{equation}

In the literature, crystal structures can be distinguished from one another by plotting the distribution of two or more $Q_{l}$.

@@

\subsection{Yield stress and zero shear rate viscosity}
\label{section:yield_stress}

@@

\subsection{Shear banding}
\label{section:yield_stress}
